---
title: "Homework 1"
author: "Lauren Alimento"
date: "3/16/2019"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
---
title: "Lab 1 Homework"
author: "Lauren Alimento"
date: "1/24/2019"
output: 
  html_document: 
    keep_md: yes
---
1. Navigate to the R console and calculate the following expressions.  
  + 5 - 3 * 2  
  + 8 / 2 ** 2  
  
```{r}
+ 5 - 3 * 2
```
```{r}
+ 8 / 2 ** 2
```

2. Did any of the results in #4 surprise you? Write two programs that calculate each expression such that the result for the first example is 4 and the second example is 16.  
The second one surprised me, because I don't know what the "**" command does. I was also pleasantly surprised when I realized the first line of code followed the order of operations.
```{r}
(+ 5 - 3)*2
```
```{r}
(+ 8 / 2)**2
```
3. You have decided to use your new analytical powers in R to become a professional gambler. Here are your winnings and losses this week.
```{r, eval=F, echo=T}
blackjack <- c(140, -20, 70, -120, 240)
roulette <- c(60, 50, 120, -300, 10)
```
a. Build a new vector called `days` for the days of the week (Monday through Friday). 
```{r, eval=F, echo=T}
days <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
days
```
We will use `days` to name the elements in the poker and roulette vectors.
```{r, eval=F, echo=T}
names(blackjack) <- days
names(roulette) <- days
```
b. Calculate how much you won or lost in blackjack over the week.
```{r, eval=F, echo=T}
total_blackjack <- sum(blackjack)
total_blackjack
```
c. Calculate how much you won or lost in roulette over the week.  
```{r, eval=F, echo=T}
total_roulette <- sum(roulette)
total_roulette
```
d. Build a `total_week` vector to show how much you lost or won on each day over the week. Which days seem lucky or unlucky for you?
```{r, eval=F, echo=T}
total_week <- c(blackjack, roulette)
total_week
```
It looks like Fridays are the luckiest days; Mondays and Wednesdays aren't bad either. Thursdays are the unluckiest days.
e. Should you stick to blackjack or roulette? Write a program that verifies this below.
```{r, eval=F, echo=T}
games <- c("Blackjack", "Roulette")
games
names(end_of_week_totals) <- games
end_of_week_totals <- c(total_blackjack, total_roulette)
end_of_week_totals
```
You should stick with blackjack. It looks like you won 310 from blackjack this week and lost 60 from Roulette this week.