---
title: "Lab 6 HW"
author: "Lauren Alimento"
date: "3/5/2019"
output: 
  html_document: 
    fig_caption: yes
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(skimr)
library("RColorBrewer")
```
```{r}
#install.packages("gapminder")
```
```{r}
library("gapminder")
```
```{r}
gapminder <- 
  gapminder::gapminder
```

1. Explore the data using the various function you have learned. Is it tidy, are there any NA's, what are its dimensions, what are the column names, etc.
```{r}
names(gapminder)
```
```{r}
gapminder
```
```{r}
glimpse(gapminder)
```

2. We are interested in the relationship between per capita GDP and life expectancy; i.e. does having more money help you live longer on average. Make a quick plot below to visualize this relationship.
```{r}
gapminder %>% 
  ggplot(aes(x=gdpPercap,y=lifeExp))+
  geom_point(size=0.25)
```

3. There is extreme disparity in per capita GDP. Rescale the x axis to make this easier to interpret. How would you characterize the relationship?

```{r}
gapminder %>% 
  ggplot(aes(x=gdpPercap,y=lifeExp))+
  geom_point(size=0.25)+
  scale_x_log10()
```

4. This should look pretty dense to you with significant overplotting. Try using a faceting approach to break this relationship down by year.

```{r}
gapminder %>%
  ggplot(aes(x=gdpPercap, y=lifeExp))+
  geom_point(size=0.25)+
  scale_x_log10()+
  facet_wrap(~year)
```

5. Simplify the comparison by comparing only 1952 and 2007. Can you come to any conclusions?
```{r}
gapminder %>%
  filter(year==1952 | year==2007) %>% 
  ggplot(aes(x=gdpPercap, y=lifeExp))+
  geom_point()+
  scale_x_log10()+
  facet_grid(~year)
```

6. Let's stick with the 1952 and 2007 comparison but make some aesthetic adjustments. First try to color by continent and adjust the size of the points by population. Add `+ scale_size(range = c(0.1, 10), guide = "none")` as a layer to clean things up a bit.
```{r}
gapminder %>%
  filter(year==1952 | year==2007) %>% 
  ggplot(aes(x=gdpPercap, y=lifeExp, color=continent, size=pop))+
  geom_point()+
  scale_x_log10()+
  facet_grid(~year)+
  scale_size(range = c(0.1, 10), guide = "none")
```

7. Although we did not introduce them in lab, ggplot has a number of built-in themes that make things easier. I like the light theme for these data, but you can see lots of options. Apply one of these to your plot above.
```{r}
?theme_light
```

8. What is the population for all countries on the Asian continent in 2007? Show this as a barplot.
```{r}
gapminder %>%
  filter(year==2007, continent=="Asia") %>% 
  ggplot(aes(x=reorder(country, -pop), y=pop))+
  geom_col(fill="steelblue")+
  theme_light()+ 
  theme(axis.text.x = element_text(angle = 60,hjust=1))
```

9. You should see that China's population is the largest with India a close second. Let's focus on China only. Make a plot that shows how population has changed over the years.
```{r}
gapminder %>%
  filter(country=="China") %>% 
  ggplot(aes(x=factor(year), y=pop))+
  geom_bar(stat="identity")
```

10. Let's compare China and India. Make a barplot that shows population growth by year using `position=dodge`. Apply a custom color theme using RColorBrewer.
```{r}
gapminder %>%
  filter(country=="China" | country=="India") %>% 
  ggplot(aes(x=factor(year), y=pop, fill=country))+
  geom_bar(stat="identity", position="dodge")+
  scale_fill_brewer(palette = "Dark2")
```

